\chapter{Keyboard and mouse controls}
\minitoc  

 \section{Keyboard and mouse controls}
\rowcolors{2}{}{gray!25}
\begin{tabularx}{\linewidth}{ | c | X | }
 \hline			
   Ctrl+A & Selects all objects \\ \hline				
   Ctrl+D & Unselects all objects \\ \hline				
   L+ left click & Creates a landmark (either ``normal", ``target" or ``flag" landmark). \\ \hline			
L + right click & If a single landmark is selected, its position is
moved. Nothing happens if no landmark is selected
or if more than one landmark are selected \\ \hline			

Left mouse button drag 
& Camera mode : camera rotation.\newline
 Object mode : object rotation.\newline
Landmark mode : camera rotation. \\ \hline			

Ctrl + left mouse button drag 
& Camera mode : object rotation.\newline
 Object mode : camera rotation.\newline
 Landmark mode : object rotation. \\ \hline	
		
Right mouse button drag & Draws a yellow rectangle. Once right button is
released, all objects (surfaces and landmarks)
falling inside the rectangle get selected/unselected,
depending on their initial selection status \\ \hline	
		
Right mouse button click & All objects (surfaces and landmarks) lying at click
location get selected/unselected, depending on
their initial selection status \\ \hline		
	
Middle mouse button roll (roll wheel) & Zoom / unzoom \\ \hline		
	
Middle mouse button drag 
& Camera mode : camera translation\newline
 Object mode : object translation\newline
 Landmark mode : camera translation \\ \hline	
		
Ctrl + middle mouse button drag 
& Camera mode : object translation\newline
Object mode : camera translation\newline
Landmark mode : object translation \\ \hline			

« Del » & All selected objects are deleted \\ \hline			

T + left click & Tag selected surface with active tag value \\ \hline			
 
T + right click & Tag selected surface with ``0" value \\ \hline			

 \end{tabularx}

\section{Additional controls}
Additional controls are available when using ``lasso cut" or ``lasso tag" (lasso mode should be active):\\
\begin{tabularx}{\linewidth}{ | c | X | }
\hline			
Left click & Adds a segment to polygon (segments are drawn yellow) \\ \hline			

Right click & Connects last segment to first segment. If two segments cross each other, lasso action is canceled. Otherwise, the closed polygon is drawn red.\\ \hline			

Middle click or ``C" + right click. & Once the lasso is closed (lasso polygon drawn red
after a right click), all the vertices falling within the clicked region (outside or inside the polygon) are either:\newline
- given the color corresponding the the active tag (lasso tag)\newline
- inserted into a new objects (lasso cut).\newline 
See lasso cut (section \ref{lasso_cut_section}) and lasso tag (section \ref{lasso_tag_section}) sections for further information.\\ \hline	
		
T + left click (``lasso tag" only) & Depending on which tag tool is active, if the
lasso is closed (lasso polygon drawn red after a right click) the vertices are painted either :\newline 
- using the pencil\newline 
- using the magic wand\newline 
- using the flood fill\newline 
Tag propagation is restricted within the area defined by the red polygon.\\ \hline		
	
\end{tabularx}

